---
title: "Integer_Optimization"
author: "Sarah Deussing"
date: "2024-12-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Integer Optimization

#### Purchasing Packages Dishes
Dish A: 2mg cholesterol, 6g fat, 6g protein
Dish B: 3mg cholesterol, 5g fat, 7g protein

Constraints <= 12mg cholesterol, <= 30g fat
  2A + 3B <= 12
  6A + 5B <= 30
  A, B >= 0

Objective: MAX(6A + 7B) --> maximize protein
```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)

obj_fun <- c(6,7)

const_LHS <- matrix(c(2, 3,
                      6, 5),
                    ncol = 2, byrow = TRUE)

const_dir <- c("<=", "<=")

const_RHS <- c(12, 30)

colnames(const_LHS) <- c("A", "B")
rownames(const_LHS) <- c("cholesterol", "fat")

int_vec <- c(1:2)

model <- lp(direction = "max",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            int.vec = int_vec)

model$objval

model$solution

cbind(const_LHS %*% model$solution, const_dir, const_RHS)
```

## Binary Integer Optimization

#### Sports Schedule
4 teams: Atlanta (A), Boston (B), Chicago (C), Detroit (D)
2 divisions: AB, CD
4 weeks: play team in division 2x, other teams 1x

Objective: MAX(preferences --> division games later)

Decision vars: x_ijk (i,j = team, k = week)

Constraints:
  x_AB1 + x_AB2 + x_AB3 + x_AB4 = 2 (same for CD)
  x_AC1 + x_AC2 + x_AC3 + x_AC4 = 1 (same for AD, BC, BD)
  x_AB1 + x_AC1 + x_AD1 = 1 (same for all combos)
```{r}
obj_fun <- c(1, 2, 4, 8, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 2, 4, 8)

# AB-1,2,3,4, AC-1,2,3,4, AD-1,2,3,4, BC-1,2,3,4, BD-1,2,3,4, CD-1,2,3,4
const_LHS <- matrix(c(1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1,
                      0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 0, 0, 0, 0,
                      1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0,
                      0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0, 0,
                      0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0, 0,
                      0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1,
                      0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 1), ncol = 24, byrow = TRUE)


const_dir <- c("=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=", "=")


const_RHS <- c(2, 2, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1, 1)


binary_vec <- c(1:22)


model <- lp(direction = "max",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            binary.vec = binary_vec)

model$objval

model$solution
```

## Facility Location
I = set of possible site distribution centers (3)
J = set of stores to be sourced (5)
  ci = cost of locating dist. center i
  fij = transportation cost from i to store j

Locate distribution centers
  xi = 1 if center at site i
  yij = 1 if store j is sourced from site i

Objective: MIN( SUM(ci`*`xi) + SUM(SUM(fij`*`yij)))

Constraints:
  SUM(yij) = 1 (each store sourced from 1 center)
  yij < xi
```{r}
# order of decision variables
# x1, x2, x3, y11, y12, ..., y15, y21, ..., y25, y31, ..., y35

obj_fun <- c(1.2, 0.8, 0.74, 0.18, 0.02, 0.28, 0.35, 0.12, 0.34, 0.21, 0.11, 0.02, 0.16, 0.39, 0.14, 0.1, 0.35, 0.37)


const_LHS <- matrix(c(0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0,
                      0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 1,
                      -1, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      -1, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      -1, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      -1, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      -1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, -1, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, -1, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0,
                      0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0,
                      0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0, 0,
                      0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0, 0,
                      0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0, 0,
                      0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0, 0,
                      0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0, 0,
                      0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0,
                      0, 0, -1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 1,
                      1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), ncol = 18, byrow = TRUE)


const_dir <- c("=", "=", "=", "=", "=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", "=")


const_RHS <- c(1, 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 2)

binary_vec <- c(1:18)

model <- lp(direction = "min",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            binary.vec = binary_vec)

model$objval

model$solution
```

