---
title: "Monte Carlo Simulation"
author: "Sarah Deussing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sum of 2 Dice

```{r}
n <-  100 # num of replications

results <-  c()

for (j in 1:n) {
  dice_outcome <-  sample(c(1,2,3,4,5,6), # outcomes
                        size = 2, # n to sample
                        replace = TRUE,
                        prob = 1/6 * c(1,1,1,1,1,1))
  sum_of_dice <-  dice_outcome[1] + dice_outcome[2]
  results <-  rbind(results, sum_of_dice)
}
```

```{r}
results_mean = mean(results)
results_std = sd(results)

# 95% confidence interval
lower_CI <-  results_mean - 1.96 * results_std/sqrt(n)
upper_CI <-  results_mean + 1.96 * results_std/sqrt(n)
lower_CI
upper_CI
```

## Investment
Given $10,000 to invest. Avg. annual return = 8% with uncertainty sd = 17%. 
Value of asset after 10 years?

X = (Y1, ..., Y10) --> return rate of year 1-10
Yi: normal distribution, mean = 1.08, sd = 0.17
f(X) = 10,000 * Y1 * ... * Y10
```{r}
n <-  100

fund_mean <-  1.08
fund_std <-  0.17
fund_initial <-  10000

results = c()

for (i in 1:n) {
  fund_value <- fund_initial
  
  for (j in 1:10) {
    return_rate <- rnorm(1, fund_mean, fund_std)
    fund_value <- fund_value * return_rate
  }
  results <- rbind(results, fund_value)
}
```

```{r}
results_mean = mean(results)
results_std = sd(results)

# 95% confidence interval
lower_CI <-  results_mean - 1.96 * results_std/sqrt(n)
upper_CI <-  results_mean + 1.96 * results_std/sqrt(n)
lower_CI
upper_CI
```

## Auto Shop Tire Profit
Demands: 300, 320, 340, 360, 380, 400
Freqs: 3, 6, 12, 18, 5, 6
Avg. selling price: `$`60 - `$`80
Avg. cost per tire ratio: 20% - 30%
Fixed cost: `$`12,000 per month

X = (D, P, V)
f(X) = Profit = P`*`D - (P`*`V)`*`D - 12000
     = P`*`D `*` (1-V) - 12000
```{r}
n = 100

results <- c()

for (i in 1:n) {
  selling_price <-  runif(1, 60, 80)
  cost_price_ratio <- runif(1, 0.2, 0.3)
  demand <- sample(c(300, 320, 340, 360, 380, 400),
                   size = 1,
                   replace = TRUE,
                   prob = 1/60 * c(3, 6, 12, 18, 5, 6))
  profit <- selling_price * demand * (1-cost_price_ratio) - 12000
  results <- rbind(results, profit)
}
```

```{r}
results_mean = mean(results)
results_std = sd(results)

# 95% confidence interval
lower_CI <-  results_mean - 1.96 * results_std/sqrt(n)
upper_CI <-  results_mean + 1.96 * results_std/sqrt(n)
lower_CI
upper_CI
```

## Inventory Management: Newsvendor Problem
Demand = normal distribution, mean = 20, sd = 4
Cost per unit = `$`5
Sale per unit = `$`25

X = uncertain demand
q = order quantity

f(X) = Profit = Revenue - Demand
     = 25`*`sales - 5`*`q
     = 25`*`min(q, X) - 5`*`q
     (sales = q if undersupply, X if oversupply)
```{r}
n = 1000

replenish_cost = 5
selling_price = 25

q_list <- c(10:30)

results <- c()

for (q in q_list) {
  for (i in 1:n) {
    demand <- rnorm(1, 20, 5)
    profit <- selling_price*min(q,demand) - replenish_cost*q
  }
  results <- rbind(results, profit)
}
```

#### With Discount
Reseller that buys at `$`4 per unit

Leftovers = inventory - sales = q - min(q, X)
New Profit = old revenue - cost + discount revenue
           = 25`*`min(q,X) - 5`*`q + 4`*`(q-min(q,X))
           
## Hotel Renovation
Standard & Luxury Rooms
Revenue: `$`99, `$`169
Cost: `$`12.5, `$`25
Supply: 60, 20

Upgrade cost: `$`50
Fixed cost: `$`6000

Demand: normal distribution (standard: mean = 50, sd = 10, luxury: mean = 20, sd = 5)

```{r}
n = 10000

revenue_standard = 99
revenue_luxury = 169

cost_standard = 12.5
cost_luxury = 25

upgrade_cost = 50
fixed_cost = 6000

simulate <- function(standard_rooms, luxury_rooms) {
  results = c()

  for (i in 1:n) {
  
    standard_dem <- rnorm(1, 50, 10)
    luxury_dem <- rnorm(1, 20, 5)
  
  #luxury_to_luxury: number of customers who originally requested luxury suites and stay there
    luxury_to_luxury <- min(luxury_dem, luxury_rooms)
  
  #standard_to_luxury: number of customers who originally requested standard rooms but upgraded to luxury suites upon availability
    standard_to_luxury <- min(standard_dem, luxury_rooms - luxury_to_luxury)
  
  #standard_to_standard: number of customers who
  # originally requested standard rooms and stay there
    standard_to_standard <- min(standard_rooms, standard_dem - standard_to_luxury)
  
    # profit
    luxury_prof <- (revenue_luxury - cost_luxury) * luxury_to_luxury
    upgraded_prof <- (revenue_standard + upgrade_cost - cost_luxury) * standard_to_luxury
    standard_prof <- (revenue_standard - cost_standard) * standard_to_standard
    profit <- luxury_prof + upgraded_prof + standard_prof - fixed_cost
  
    results = rbind(results, profit)
    }
  results_mean = mean(results)
  results_sd = sd(results)
  lower_CI = results_mean - 1.96 * results_sd / sqrt(n)
  upper_CI = results_mean + 1.96 * results_sd / sqrt(n)
  print(lower_CI)
  print(upper_CI)
}

set.seed(1)
simulate(60, 20)
simulate(40, 30)

# The plan with 40 standard rooms and 30 luxury rooms is the better option because it 
# has higher values for the confidence interval, i.e. a greater profit
```

