---
title: "PS1_Deussing"
author: "Sarah Deussing"
date: "2024-11-10"
output: html_document
---

## 1: Investment Example Revisited
```{r}
set.seed(1)
```


```{r}
# Number of scenarios
n = 1000

# Parameters
fund_mean = 1.08
fund_std = 0.17
fund_initial_value = 10000

# Cash conversion rate
#cash_rate = 50

# Run simulation model
simulate <- function(cash_rate) {

results = c()

for (i in 1:n) {
  
  # Initialize fund_value and cash_value (amount of cash)
  fund_value = fund_initial_value
  cash_value = 0
  
  for (j in 1:3) {
    # TO-DO: Randomly generate scenarios of annual return rate
    return_rate = rnorm(1, mean = fund_mean, sd = fund_std)
    
    # TO-DO: Update fund_value and cash_value accordingly
    fund_value = fund_value * return_rate
    cash_value = cash_value + (cash_rate/100 * fund_value)
    fund_value = fund_value - cash_value
  }
  
  # After three years, remaining funds are converted to cash
  total_value = cash_value + fund_value
  
  # Add to results
  results = rbind(results, total_value)
}

# Compute mean and standard deviation
results_mean = mean(results)
results_std = sd(results)

cat("Cash_rate:", cash_rate, "%\n")
cat("Mean:", results_mean, "\n")
cat("Sd:", results_std, "\n")

# Compute 95% confidence interval
lower_CI = results_mean - 1.96 * results_std / sqrt(n)
upper_CI = results_mean + 1.96 * results_std / sqrt(n)
cat("Confidence interval:", lower_CI, ",", upper_CI, "\n\n")
}

simulate(cash_rate = 50)
simulate(cash_rate = 0)
```

The confidence interval for a cash rate of 50% is significantly higher than when the cash rate is 0%. This means that the cash rate of 50% will generate a higher total cash generated from the initial investment, so this would be the preferred r-value.

## 2: Quality Improvement
```{r}
# Number of scenarios
n = 10000

# Investment decision
#investment = 5000

# Run simulation model
simulate2 = function(investment) {

results = c()

for (i in 1:n) {
  
  # TO-DO: Randomly generate scenarios of selling price (P) and demand (Q)
  selling_price = runif(1, min = 1+(investment/10000), max = (1.15+(1.5*investment/10000)))
  demand = rnorm(1, mean = (10000-(.1*investment)), sd = 1000)
  
  # TO-DO: Express profit as a function of Q, P and investment
  profit = (demand*selling_price) - investment
  
  # Add to results
  results = rbind(results, profit)
}

# Compute mean and standard deviation
results_mean = mean(results)
results_std = sd(results)

# Compute 95% confidence interval
lower_CI = results_mean - 1.96 * results_std / sqrt(n)
upper_CI = results_mean + 1.96 * results_std / sqrt(n)

cat("Investment:", investment, "\n")
cat("Confidence interval:", lower_CI, ",", upper_CI, "\n\n")
}

simulate2(investment = 0)
simulate2(investment = 5000)
simulate2(investment = 15000)
```

The highest values for the confidence interval occur when the investment is `$`5,000. This means that investing `$`5,000 will yield the greatest profit for the company, so they should select this option.

## 3: Inventory Management with Clearance
```{r}
# Number of scenarios
n = 10000


# Parameters
wholesale_price = 11
retail_price = 16.95
clearance_price = 6.95

# Ordering quantity
#Q = 3500

# Run simulation model
simulate3 = function(Q) {

results = c()

for (i in 1:n) {
  # TO-DO: Randomly generate a scenario of demand
  demand = runif(1, 1000, 5000)
  
  
  # TO-DO: Express potential leftover by using Q and D
  leftover = max(Q - demand,0)
  
  # TO-DO: Compute profit
  profit = min(demand, Q) * (retail_price - wholesale_price) + (leftover*(clearance_price - wholesale_price))
  
  # Add to results
  results = rbind(results, profit)
}

# Compute mean and standard deviation
results_mean = mean(results)
results_std = sd(results)

# Compute 95% confidence interval
lower_CI = results_mean - 1.96 * results_std / sqrt(n)
upper_CI = results_mean + 1.96 * results_std / sqrt(n)
cat("Q:", Q, ", ")
cat("Confidence interval:", lower_CI, ",", upper_CI, "\n")
return(c(lower_CI = lower_CI, upper_CI = upper_CI))
}


order_quantities = seq(3000,4000, by = 50) #simulate for numbers 1000-6000 (every 50th number)
CIs = sapply(order_quantities, simulate3)

colnames(CIs) = order_quantities
CIs = t(CIs)
CIs = as.data.frame(CIs)

best_row = which.max(CIs$upper_CI) #find interval with highest values for greatest profit
CIs[best_row,]
```

The highest values for the confidence interval are when Q = 5,000 units, so this should be the number of units purchased.

Testing with new retail price
```{r}
# New Parameters
wholesale_price = 11
retail_price = 18.95
clearance_price = 6.95

# Run the same simulation with these new parameters
CIs_new = sapply(order_quantities, simulate3)

colnames(CIs_new) = order_quantities
CIs_new = t(CIs_new)
CIs_new = as.data.frame(CIs_new)

best_row_new = which.max(CIs_new$upper_CI)

# Compare confidence interval values
CIs[best_row,]
CIs_new[best_row_new,]
```

The best order quantity, Q, remains the same at 5000 units. However, the values of the confidence interval are much higher, meaning that there will be a greater profit seen with this new retail price. This is true across all order quantities, the values of the new confidence intervals are higher than those of the original.

