---
title: "PS2_Deussing"
author: "Sarah Deussing"
date: "2024-11-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## (1) Zippy Lube
Part A
```{r}
hours_per_week = 60 # 10 hours/day * 6 days/week
profit_per_oc = 15 #per oil change
lambda = 3.5 #cars/hour = arrival rate
mieu = 4.3 #cars/hour = service rate

W = 1/(mieu - lambda)
L = lambda/(mieu - lambda)
Wq = lambda/((mieu - lambda)*mieu)
Lq = (lambda*lambda)/((mieu - lambda)*mieu)

paste0("The average number of vehicles waiting for an oil change = ", Lq)
```

Part B
```{r}
oc_per_week = lambda * hours_per_week
profit = oc_per_week * profit_per_oc
paste("The profit per week = $", profit)
```

Part C
```{r}
current_ser_time = 1/mieu * 60 # in minutes
new_ser_time = current_ser_time - 1
new_ser_time = new_ser_time / 60 # back to hours

new_mieu = 1/new_ser_time
new_profit = new_mieu * hours_per_week * profit_per_oc
add_profit = new_profit - profit

payback_period = 5000/add_profit
paste0("The payback period is ", payback_period, " weeks.")
```

## (2) Amazon Hub Lockers
Part A
```{r}
# Load the package
library(queueing)

# Define parameters (arrival and service rates and number of servers)
c = 50 # num hub lockers
paste0("c: ",c)
lambda = 120/24 #120 per day / 24 = arrival rate per hour
paste0("lambda: ",lambda)
mieu = 1/8 # 1/8 hours service time = service rate
paste0("mieu: ",mieu)
```

The "servers" are the number of hub lockers (in this example, 50).
The "service time" is the time it takes for a customer to pick up their package (in this example, 8 hours). The "service rate" is 1/service time --> mieu.
The "arrival rate" is the number of package requests per day (in this example, 120). --> lambda
  - We need to convert this to an hourly value to match the unit for the service time. 

```{r}
# Use the functions within queueing package to create your queueing model
amazonModel <- NewInput.MMC(lambda = lambda,
                            mu = mieu,
                            c = c)
amazonQueue <- QueueingModel(amazonModel)

# Print relevant operating characteristics to conclude
avg_delay = (amazonQueue$Wq) * 60 #Wq = average wait time in queue
paste0("The average delay is: ",avg_delay, " mins.")
```

Part C
```{r}
new_lambda = 145/24 # increased requests to 145/day
max_delay = 1 # 1 hour

original_c <- c
while(TRUE) {
  amazonModel_c <- NewInput.MMC(lambda = new_lambda,
                            mu = mieu,
                            c = c)
  amazonQueue_c <- QueueingModel(amazonModel_c)
  avg_delay_c <- amazonQueue_c$Wq
  if (avg_delay_c <= max_delay) { break }
  c <- c+1
}
add_lockers <- c - original_c
paste0("The number of additional lockers needed is: ", add_lockers)
```


## (3) TSA Pre-Check
Part A + B
```{r}
# Load the package
library(queueing)

# Define parameters (arrival and service rates, number of servers, etc.)
arrival_rate = 115 # per hour, for both = lambda_total
service_rate_regular = 60/5 # 5 minutes per person
service_rate_precheck = 60/3 # 3 minutes per person

# Number of total workers
c_overall = 10

# Proportion of customers enrolled in TSA PreCheck
precheck_rate = 0.16 # 16% of people
lambda_regular = arrival_rate*(1-precheck_rate)
lambda_precheck = arrival_rate*precheck_rate

# Create the input for the model.
optimize <- function(lambda, mu, c) {
  rho <- lambda / (mu * c)
  if (rho >= 1) { return(NULL) }
  
  airportModel <- NewInput.MMC(lambda = lambda, mu = mu, c = c)
  airportQueue <- QueueingModel(airportModel)
  return(airportQueue$Lq)
}

# Create the queueing model for each of the three
min_wait_time <- Inf
best_servers <- c(0,0) #n. precheck, n. regular servers

for (c_precheck in 0:9){
  c_regular = c_overall - c_precheck
  
  Lq_precheck <- optimize(lambda_precheck,
                          service_rate_precheck,
                          c_precheck)
  Lq_regular <- optimize(lambda_regular,
                          service_rate_regular,
                          c_regular)
  if (is.null(Lq_precheck) || is.null(Lq_regular)) {next}
  
  Lq_total <- Lq_precheck + Lq_regular
  print(paste0("With ", c_precheck, " PreCheck server(s) and ", c_regular, " Regular server(s), total wait: ", Lq_total))
  
  if(Lq_total < min_wait_time) {
    min_wait_time <- Lq_total
    best_servers <- c(c_precheck, c_regular)
  }
}

# Print relevant operating characteristics and compare
paste0("The best allocation is: Precheck = ", best_servers[1], " servers and Regular = ", best_servers[2], " servers.")

```
The optimal assignment of servers is: 1 precheck and 1 regular.

Part C
```{r}
best_servers_tot <- list()
total_wait_times <- c()

init_precheck_rate = 0.16
num_periods = 6 # 3 years * 6 months

for (period in 2:num_periods) {
  precheck_rate = init_precheck_rate + 0.04 * (period-1)
  lambda_regular = arrival_rate * (1 - precheck_rate)
  lambda_precheck = arrival_rate * precheck_rate
  
  min_time <- Inf
  best_servers <- c(0,0)
  
  for (c_precheck in 1:9) {
    c_regular = c_overall - c_precheck
    
    Lq_precheck <- optimize(lambda_precheck,
                            service_rate_precheck,
                            c_precheck)
    Lq_regular <- optimize(lambda_regular,
                           service_rate_regular,
                           c_regular)
    only_regular <- optimize(115, service_rate_regular, 10)
    #if (is.null(Lq_precheck) || is.null(Lq_regular)) {next}
    
    Lq_total <- Lq_precheck + Lq_regular
    best <- min(Lq_total, only_regular)
    if (best < min_wait_time) {
      min_time <- best
      if (best == only_regular) { best_servers <- c(0,10)}
      else {best_servers <- c(c_precheck, c_regular)}
    }
  }
  best_servers_tot[[period]] <- best_servers
  total_wait_times[period] <- min_wait_time
  
  print(paste0("Period ", period, ",PreCheck rate: ",precheck_rate))
  print(paste0("Best allocation: PreCheck = ", best_servers[1], " servers, Regular = ", best_servers[2], " servers"))
  print(paste0("  Total waiting time: ", min_time))
}


print("Long-term Server Allocation Plan for the Next 3 Years:")
for (period in 2:num_periods) {
  print(paste0("Period ", period,
         ": PreCheck = ", best_servers_tot[[period]][1],
         " servers, ",
         "Regular = ", best_servers_tot[[period]][2],
         " servers"))
}
```

