---
title: "PS3"
author: "Sarah Deussing"
date: "2024-12-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1: Optimizing Marketing Research Schedule
Objective: minimize cost
15HwC1 + 18HwC2 + 13HwoC1 + 16HwoC2 + 10OH1 + 12OH2

Constraints:
1. Households w/ children >= 400
  - >= 50% at night --> −1HwC1 + 1HwC2 >= 0
  - cost: 15 for day and 18 for night
2. Households w/o children >= 200
  - >= 60% at night --> -1.5HwoC1 + 1HwoC2 >= 0
  - cost: 13 for day and 16 for night
3. One-person households >= 100
  - >= 70% at night --> -(7/3)OH1 + 1OH2 >= 0
  - cost: 10 for day and 12 for night

total interviews = 800 (# night > # day)

Decision variables: HwC, HwoC, OH --> 1 = day, 2 = night

HwC1 + HwC2 + HwoC1 + HwoC2 + OH1 + OH2 <= 800
HwC1 + HwC2 >= 400
HwoC1 + HwoC2 >= 200
OH1 + OH2 >= 100

```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)

obj_fun <- c(15, 18, 13, 16, 10, 12)

# left-hand side vector
const_LHS <- matrix(c(1, 1, 1, 1, 1, 1, # total
                      1, 1, 0, 0, 0, 0, # total w/ children
                      0, 0, 1, 1, 0, 0, # total w/o children
                      0, 0, 0, 0, 1, 1, # total one-person
                      -1, 1, 0, 0, 0, 0, # night > day
                      0, 0, -1.5, 1, 0, 0, # night > day
                      0, 0, 0, 0, -(7/3), 1), # night > day
                    ncol = 6, byrow = TRUE)


const_dir <- c(">=", ">=", ">=", ">=", ">=", ">=", ">=")

colnames(const_LHS) = c("HwC1", "HwC2", "HwoC1", "HwoC2", "OH1", "OH2")
rownames(const_LHS) = c("total", "total_w_children", "total_wo_children", "total_one_person",
                        "w_children_night", "wo_children_night", "one_person_night")

# right-hand side vector
const_RHS <- c(800, 400, 200, 100, 0, 0, 0)

# model
model1 <- lp(direction = "min",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            compute.sens = TRUE)

model1
model1$objval
model1$solution

cbind(const_LHS %*% model1$solution, const_dir, const_RHS)
```

Lowest price: 11840
Total w/ children day: 200
Total w/ children night: 200
Total w/o children day: 80
Total w/o children night: 120
Total 1 person day: 60
Total 1 person night: 140

## 2: Investment Management Under Taxation
Objective: maximize profit (goal of 10,000)
150 shares of 8 different stocks --> 1,200 shares

- 1% transaction cost for selling
- 30% capital gains tax at time of sale

#### Part A:
(i) Decision Variables: # of shares to sell
 - x1 (Yahoo), x2 (GE), x3 (Microsoft), x4 (BoA), x5 (JPMC), x6 (Cisco), x7 (Intel), x8 (Pfizer)
 
(ii) Constraint
 - cannot sell negative stock (0 <= x)
 - net cash per share: 0.99current price - 0.3purchase_price
 

(iii) Estimated value of Microsoft stock
 - (150 - x)* $34.55 = value = 150(34.55) - 34.55x

(iv) Objective Function: Max value of portfolio next year
29.5(150-x1) + 26.31(150-x2) + 34.55(150-x3) + 15.23(150-x4) + 62.43(150-x5) + 26.68(150-x6) + 23.85(150-x7) + 31.66(150-x8)
This is equal to: (150-xi)next_year_price = 150(next_year_price) - xi(next-year_price). So the objective function is -next_year_price and we have to subtract the objective value from 150(next_year_price) at the end.


(v) $10,000 Constraint
Net price gain >= 10,000

net_price_gain <- 0.99(xi)current_price -  0.3((xi)current_price-150*purchase_price)
 = 0.99(x)curr  - (0.3(x)curr - 0.3(150)purchase)
 = 0.99(x)curr - 0.3(x)curr  + 0.3(150)purchase
 = x(0.99curr - 0.3curr) + 0.3(150)purchase

So we have net_price_gain = x(0.99curr - 0.3curr)
and the constraint is: 10000 - 0.3(150)purchase

#### Part B:
```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)



purchase_price <- c(15.68, 22.10, 30.39, 8.93, 40.55, 18.58, 22.54, 24.84)
current_price <- c(31.80, 24.28, 32.50, 14.16, 50.99, 24.17, 23.67, 28.77)
next_year_prices <- c(29.50, 26.31, 34.55, 15.23, 62.43, 26.68, 23.85, 31.66)

obj_fun <- -next_year_prices # (150-x) * price for next year

# calculate net_price_gain
# = 0.99*xi*current_price -  0.3*(xi*current_price-150*purchase_price) # pull out xi
# = 0.99*x*curr  - (0.3*x*curr - 0.3*150*purchase)
# = 0.99*x*curr - 0.3*x*curr  + 0.3*150*purchase
# = x(0.99*curr - 0.3*curr) + 0.3*150*purchase
net_price_gain <- 0.99*current_price -  0.3*current_price # then add 0.3*150*purchase


# left-hand side vector
const_LHS <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, # yahoo
                      0, 1, 0, 0, 0, 0, 0, 0, # GE
                      0, 0, 1, 0, 0, 0, 0, 0, # microsoft
                      0, 0, 0, 1, 0, 0, 0, 0, # BofA
                      0, 0, 0, 0, 1, 0, 0, 0, # JPMC
                      0, 0, 0, 0, 0, 1, 0, 0, # cisco
                      0, 0, 0, 0, 0, 0, 1, 0, # intel
                      0, 0, 0, 0, 0, 0, 0, 1, # pfizer
                      1, 0, 0, 0, 0, 0, 0, 0, # yahoo
                      0, 1, 0, 0, 0, 0, 0, 0, # GE
                      0, 0, 1, 0, 0, 0, 0, 0, # microsoft
                      0, 0, 0, 1, 0, 0, 0, 0, # BofA
                      0, 0, 0, 0, 1, 0, 0, 0, # JPMC
                      0, 0, 0, 0, 0, 1, 0, 0, # cisco
                      0, 0, 0, 0, 0, 0, 1, 0, # intel
                      0, 0, 0, 0, 0, 0, 0, 1, # pfizer
                      net_price_gain), 
                    ncol = 8, byrow = TRUE)


const_dir <- c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", ">=", ">=", ">=", ">=", ">=", ">=", ">=", ">=", ">=")

colnames(const_LHS) = c("x1", "x2", "x3", "x4", "x5", "x6", "x7", "x8")
rownames(const_LHS) = c("Yahoo", "GE", "Microsoft", "BofA", "JPMC", "Cisco", "Intel", "Pfizer", "Yahoo", "GE", "Microsoft", "BofA", "JPMC", "Cisco", "Intel", "Pfizer", "price_min")

# right-hand side vector
const_RHS <- c(150, 150, 150, 150, 150, 150, 150, 150, 
               0, 0, 0, 0, 0, 0, 0, 0, (10000 - sum(0.3*150*purchase_price)))

# model
model2 <- lp(direction = "max",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            compute.sens = TRUE)

model2

sum(150*next_year_prices) - model2$objval # (150 - x) * next = 150*next - x*next
model2$solution

cbind(const_LHS %*% model2$solution, const_dir, const_RHS)
```

(i) Optimal solution: only sell 79.18831 of Yahoo stock for a net price gain of 39867.56
(ii) An investor that would prefer this solution is someone who wants to sell a significant amount of only one stock to have a greater focus on the remaining companies. 

#### Part C:
- keep >= 75 shares of each stock --> can only sell 75
```{r}
const_LHS <- matrix(c(1, 0, 0, 0, 0, 0, 0, 0, # yahoo
                      0, 1, 0, 0, 0, 0, 0, 0, # GE
                      0, 0, 1, 0, 0, 0, 0, 0, # microsoft
                      0, 0, 0, 1, 0, 0, 0, 0, # BofA
                      0, 0, 0, 0, 1, 0, 0, 0, # JPMC
                      0, 0, 0, 0, 0, 1, 0, 0, # cisco
                      0, 0, 0, 0, 0, 0, 1, 0, # intel
                      0, 0, 0, 0, 0, 0, 0, 1, # pfizer
                      1, 0, 0, 0, 0, 0, 0, 0, # yahoo
                      0, 1, 0, 0, 0, 0, 0, 0, # GE
                      0, 0, 1, 0, 0, 0, 0, 0, # microsoft
                      0, 0, 0, 1, 0, 0, 0, 0, # BofA
                      0, 0, 0, 0, 1, 0, 0, 0, # JPMC
                      0, 0, 0, 0, 0, 1, 0, 0, # cisco
                      0, 0, 0, 0, 0, 0, 1, 0, # intel
                      0, 0, 0, 0, 0, 0, 0, 1, # pfizer
                      net_price_gain), 
                    ncol = 8, byrow = TRUE)


const_dir <- c("<=", "<=", "<=", "<=", "<=", "<=", "<=", "<=", ">=", ">=", ">=", ">=", ">=", ">=", ">=", ">=", ">=")

colnames(const_LHS) = c("x1", "x2", "x3", "x4", "x5", "x6", "x7", "x8")
rownames(const_LHS) = c("Yahoo", "GE", "Microsoft", "BofA", "JPMC", "Cisco", "Intel", "Pfizer", "Yahoo", "GE", "Microsoft", "BofA", "JPMC", "Cisco", "Intel", "Pfizer", "price_min")

# right-hand side vector
const_RHS <- c(75, 75, 75, 75, 75, 75, 75, 75, # changed from 150 in Part B
               0, 0, 0, 0, 0, 0, 0, 0, (10000 - sum(0.3*150*purchase_price)))

# model
model3 <- lp(direction = "max",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            compute.sens = TRUE)

model3

sum(150*next_year_prices) - model3$objval
model3$solution
cbind(const_LHS %*% model3$solution, const_dir, const_RHS)
```

(i) Optimal solution: sell 75 of Yahoo and 5.626887 of Intel. For this problem, only the last constraint changed, the decision variables and outcome function remained the same. 
(ii) The investor that would prefer this solution is someone that wants to keep a diversified stock (i.e. not sell all of any stocks). This individual is concerned about the value of their stock next year because they are selling less of one specific stock to get a slightly higher gain in next year's value.

## 3: Production Planning
Outcome: Minimize total production cost
  - 20A + 25B
Bike A:
  - requires 1.5M + 1.1E 
  - cost: 20
  - max production per month: 1000
Bike B:
  - 1.6M + 1.2E
  - cost: 25
  - max production per month: 1000
  
Currently: 100 of A, 120 of B in stock
At end: need to have >= 130 of A, >= 110 of B in stock

#### Part A:
Decision vars:
- x1 (A May), x2 (A June), x3 (A July), y1 (B May), y2 (B June), y3 (B July)

Constraints:
- Max Production:
   - all vars <= 1000
- Min Demand:
   - A1 >= 700, B1 >= 800, A2 >= 900, B2 >= 600, A3 >= 1000, B3 >= 900
- Max Capacity (Machine + Employee):
   - May: 1.5(x1) + 1.6x2 <= 3000, 1.1(x1) + 1.2(y1) <= 2500
   - June: 1.5(x1) + 1.6x2 <= 2800, 1.1(x1) + 1.2(y1) <= 2300
   - July: 1.5(x1) + 1.6x2 <= 3600, 1.1(x1) + 1.2(y1) <= 2400
- Stock:
   - A1 = 100 + x1 >= 700 --> x1 >= 600
   - B1 = 120 + y1 >= 800 --> y1 >= 680
   - A2 = A1 + x2 >= 900 --> (100 + x1) + x2 >= 900 --> x1 + x2 >= 800
   - B2 = B1 + y2 >= 600 --> (120 + y1) + y2 >= 600 --> y1 + y2 >= 480
   - A3 = A2 + x3 >= 1000 + 130 --> (100 + x1 + x2) + x3 >= 1130 --> x1 + x2 + x3 >= 1030
   - B3 = B2 + y3 >= 900 + 110 --> (120 + y1 + y2) + y3 >= 1010 --> y1 + y2 + y3 >= 890


#### Part B:
```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)

obj_fun <- c(20, 25, 20, 25, 20, 25) # production costs

# left-hand side vector
const_LHS <- matrix(c(1, 0, 0, 0, 0, 0, # all vars max production
                      0, 1, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0,
                      0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 1,
                      1, 0, 0, 0, 0, 0, # all vars stock equations
                      0, 1, 0, 0, 0, 0,
                      1, 0, 1, 0, 0, 0,
                      0, 1, 0, 1, 0, 0,
                      1, 0, 1, 0, 1, 0,
                      0, 1, 0, 1, 0, 1,
                      1.5, 1.6, 0, 0, 0, 0, # machine max
                      1.1, 1.2, 0, 0, 0, 0, # employee max
                      0, 0, 1.5, 1.6, 0, 0,
                      0, 0, 1.1, 1.2, 0, 0,
                      0, 0, 0, 0, 1.5, 1.6,
                      0, 0, 0, 0, 1.1, 1.2),
                    ncol = 6, byrow = TRUE)


const_dir <- c("<=", "<=", "<=", "<=", "<=", "<=",
               ">=", ">=", ">=", ">=", ">=", ">=",
               "<=", "<=", "<=", "<=", "<=", "<=")

colnames(const_LHS) = c("x1", "y1", "x2", "y2", "x3", "y3")
rownames(const_LHS) = c("Max_x1", "Max_y1", "Max_x2", "Max_y2", "Max_x3", "Max_y3", 
                        "A1", "B1", "A2", "B2", "A3", "B3",
                        "Machine_May", "Emp_May", "Machine_June", "Emp_June", "Machine_July", "Emp_July")

# right-hand side vector
const_RHS <- c(1000, 1000, 1000, 1000, 1000, 1000,
               600, 680, 800, 480, 1030, 890,
               3000, 2500, 2800, 2300, 3600, 2400)

# model
model4 <- lp(direction = "min",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            compute.sens = TRUE)

model4

model4$objval
model4$solution
# constraint satisfaction
cbind(const_LHS %*% model4$solution, const_dir, const_RHS)
```
#### Part C:
The optimal production planning schedule states that bike A is only produced during May and June (1000 in May + 30 in June), and bike B is only produced in May (890 in May). Afterwards, the optimal solutions states that the company will rely only on leftovers.

#### Part D:
```{r}
duals <- t(model3$duals[1:18])
t(data.frame(rbind(rownames(const_LHS), duals)))

```

Constraints A3 and B3 which are the lower bounds for stock required for bikes A and B at the end of July are non-zero shadow prices. This means that for every one decrease in the lower bound for stock in July of A, the cost can be minimized by another \$20 and for every one decrease in the lower bound for stock of July of B, the cost can be minimized by another \$25.

#### Part E:

In stock at the end of May for bike A: 1000
In stock at the end of June for bike A: 1030
In stock at the end of May for bike B: 890
In stock at the end of June for bike B: 890



#### Part F:

Storage May: x1 + y1 - 700 - 800 <= 400 --> x1 +y1 <= 1900
Storage Jun: x1 + y1 - 700 - 800 + x2 + y2 -900 -600 <= 400 --> x1 +y1 x2 +y2<= 2600
```{r}
library(lpSolve, quietly = TRUE)
library(kableExtra, quietly = TRUE)

obj_fun <- c(20, 25, 20, 25, 20, 25) # production costs

# left-hand side vector
const_LHS <- matrix(c(1, 0, 0, 0, 0, 0, # all vars max production
                      0, 1, 0, 0, 0, 0,
                      0, 0, 1, 0, 0, 0,
                      0, 0, 0, 1, 0, 0,
                      0, 0, 0, 0, 1, 0,
                      0, 0, 0, 0, 0, 1,
                      1, 0, 0, 0, 0, 0, # all vars stock equations
                      0, 1, 0, 0, 0, 0,
                      1, 0, 1, 0, 0, 0,
                      0, 1, 0, 1, 0, 0,
                      1, 0, 1, 0, 1, 0,
                      0, 1, 0, 1, 0, 1,
                      1, 1, 0, 0, 0, 0, #storage may
                      1, 1, 1, 1, 0, 0, #storage jun
                      1.5, 1.6, 0, 0, 0, 0, # machine max
                      1.1, 1.2, 0, 0, 0, 0, # employee max
                      0, 0, 1.5, 1.6, 0, 0,
                      0, 0, 1.1, 1.2, 0, 0,
                      0, 0, 0, 0, 1.5, 1.6,
                      0, 0, 0, 0, 1.1, 1.2),
                    ncol = 6, byrow = TRUE)


const_dir <- c("<=", "<=", "<=", "<=", "<=", "<=",
               ">=", ">=", ">=", ">=", ">=", ">=",
               "<=", "<=",
               "<=", "<=", "<=", "<=", "<=", "<=")

colnames(const_LHS) = c("x1", "y1", "x2", "y2", "x3", "y3")
rownames(const_LHS) = c("Max_x1", "Max_y1", "Max_x2", "Max_y2", "Max_x3", "Max_y3", 
                        "A1", "B1", "A2", "B2", "A3", "B3", "Storage_May", "Storage_June",
                        "Machine_May", "Emp_May", "Machine_June", "Emp_June", "Machine_July", "Emp_July")

# right-hand side vector
const_RHS <- c(1000, 1000, 1000, 1000, 1000, 1000,
               600, 680, 800, 480, 1030, 890, 1900, 2600,
               3000, 2500, 2800, 2300, 3600, 2400)

# model
model5 <- lp(direction = "min",
            objective.in = obj_fun, 
            const.mat = const_LHS,
            const.dir = const_dir,
            const.rhs = const_RHS,
            compute.sens = TRUE)

model5

model5$objval
model5$solution
# constraint satisfaction
cbind(const_LHS %*% model5$solution, const_dir, const_RHS)
```

The storage constraints change the solution so that for bike A, 800 are produced in May, 0 are produced in June and 230 are produced in July and for bike B, 890 are produced in May, 0 are produced in June and 0 are produced in July.