---
title: "QueueingModels_Simulation"
author: "Sarah Deussing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## M/M/1 Model
1 server

#### Muffler Shop 
Mechanic can replace at avg. rate 3 per hour
Customers arrive every 30 mins on average

lambda = 1/30 mins = 1/0.5 hr = 2/hour
mu = 3/hour

W (time in system) = 1/mu - lambda = 1 hour

Cost of cust. waiting = `$`10 per hr per customer
Mechanic wage = `$`15 per hour
Purchasing cost = `$`200 per unit

Total cost/hour = customer + wage + purchasing
                = `$`10`*`L + `$`15 + `$`200`*`lambda
                = `$`20 + `$`15 + `$`400 = `$`435
To break even, charge `$`435/2 = `$`217.5 /customer

## M/M/c Model
c servers

#### Customer Support Hotline
6 Technicians, each services calls at rate of 4/hour
Calls arrive randomly at rate of 20/hour

Number of technicians to make wait time < 2 minutes?
  --> Wq

lambda = 20, mu = 4, c = 6
```{r}
library(queueing)

hotlineModel <- NewInput.MMC(20, 4, 8)
hotlineQueue <- QueueingModel(hotlineModel)

print(summary(hotlineQueue), digits = 3)

print(paste0('Wait time in mins: ', 60 * hotlineQueue$Wq))
```

## M/M/c/k Models
Finite queue length

#### Example structure
```{r, warning = FALSE}
library(queueing)

k = 6

MMCK_Model <- NewInput.MMCK(5, 7, 1, k)
MMCK_queue <- QueueingModel(MMCK_Model)

print(summary(MMCK_queue))

# distribution of # customers in the system
MMCK_queue$Pn

# balking probability
balk_prob <- MMCK_queue$Pn[k+1]
balk_prob
```

#### Gas Station Waiting Space
2 gas pumps, 3-vehicle waiting area
Vehicles arrive at rate of 28/hour
Vehicle spends 4 mins on average to fuel = 

Average profit per customer = `$`30
To expand waiting area = `$`12 /hour /space

lambda = 28, mu = 15, s = 2, q - 3, k = q+s

Total cost = rental + opportunity
           = `$`12`*`q + `$`30`*`lambda`*`prob(balk)
```{r}
library(queueing)

results <- c()

for (q in 3:9) { #testing out different wait capacity
  k = 2 + q
  
  gasModel <- NewInput.MMC(28, 15, 2, k)
  gasQueue <- QueueingModel(gasModel)
  balk_prob <- gasQueue$Pn[k+1]
  
  land_cost <- 12*q
  opp_cost <- 30*28*balk_prob
  total_cost <- land_cost + opp_cost
  
  results <- rbind(results, cbind(q, total_cost))
}
results
```

#### Hospital Capacity
Big hospital: lambda = 14.1 /day, LOS = 2.9 days, c = 56 beds
Small hospital: lambda = 4/8 /day, LOS = 2.9 days, c = 19 beds

Probability of delay?
```{r}
library(queueing)

bigModel <- NewInput.MMC(14.1, 1/2.9, 56)
smallModel <- NewInput.MMC(4.8, 1/2.9, 19)

bigQueue <- QueueingModel(bigModel)
smallQueue <- QueueingModel(smallModel)

# Occupancy Levels
print(paste("Big: ", bigQueue$RO))
print(paste("Small: ", smallQueue$RO))

# Probability of Delay
print(paste("Big: ", 1 - bigQueue$FWq(0)))
print(paste("Small: ", 1 - smallQueue$FWq(0)))
```


